Optimizing a PoS Tag Set for Dependency Parsing

### 2016-05-03 submission.tex
Additions to previous work, small fixes

\section{Introduction}
\label{sec:intro}
PoS tagging is an important preprocessing step for many NLP
tasks, such as parsing, named entity recognition and sentiment
analysis. Whereas much effort has gone into the development of
PoS-taggers, to the effect that this task is often considered
more or less a solved task, considerably less effort has been
devoted to the empirical evaluation of the PoS tag sets
themselves.
PoS tag sets are usually employed in a "one size fits
all" fashion, regardless of the requirements posed by the
downstream NLP tasks for which PoS-tagging is performed.

It is well known that syntactic parsing often benefits from quite
fine-grained morphological distinctions. However, in a realistic
setting where the aim is to parse raw text, this will often
require a separate step of morphological analysis in addition to
PoS-tagging that can be quite expensive.

In this paper, we optimize a Norwegian PoS tag set for the task of
data-driven dependency parsing.  We show that the introduction of
morphological distinctions not present in the original tag set, whilst
compromising tagger accuracy, actually leads to significantly improved
parsing accuracy.

The article is structured as follows. We start out by reviewing
previous work on tag set evaluation in Section \ref{sec:prev} and

>>>>

\section{Introduction}
\label{sec:intro}
PoS tagging is an important preprocessing step for many NLP
tasks, such as parsing, named entity recognition and sentiment
analysis. Whereas much effort has gone into the development of
PoS-taggers, to the effect that this task is often considered
more or less a solved task, considerably less effort has been
devoted to the empirical evaluation of the PoS tag sets
themselves.
PoS tag sets are usually employed in a "one size fits
all" fashion, regardless of the requirements posed by the
devoted to empirical evaluation of the PoS tag sets
themselves.
PoS tag sets are usually employed in a ``one size fits
all'' fashion, regardless of the requirements posed by the
downstream NLP tasks for which PoS-tagging is performed.

It is well known that syntactic parsing often benefits from quite
fine-grained morphological distinctions since morphology interacts
with syntax though phenomena such as agreement and case
marking. However, in a realistic setting where the aim is to
automatically parse raw text, the generation of morphological
information will often require a separate step of morphological
analysis that can be quite costly.

In this paper, we optimize a Norwegian PoS tag set for the task of
data-driven dependency parsing.  We show that the introduction of
morphological distinctions not present in the original tag set, whilst
compromising tagger accuracy, actually leads to significantly improved
parsing accuracy.
data-driven dependency parsing. In a set of experiments, various
morphological distinctions are introduced and evaluated both
intrinsically, i.e. in terms of PoS-tagging accuracy and
extrinsically, in terms of parsing accuracy.  Our results show that
the introduction of morphological distinctions not present in the
original tag set, whilst compromising tagger accuracy, actually leads
to significantly improved parsing accuracy. This optimization allows
us to bypass the additional step of morphological analysis, framing
the whole preprocessing problem as a simple tagging task.

The article is structured as follows. We start out by reviewing
previous work on tag set evaluation in Section \ref{sec:prev} and
\ref{sec:data} provides an overview of the treebank used for
experimentation. Section \ref{sec:setup} describes the experimental
setup used in our work and Section \ref{sec:optimization} goes on to
provide the results from our optimization experiments. Finally,
Section \ref{sec:summary} summarizes our main findings and discusses
some avenues for future work.
